HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-06-22 scnn.tex
commit before Lausanne

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.95\textwidth]{figures/smooth_map_class_1.pdf}
\includegraphics[width=0.95\textwidth]{figures/smooth_map_class_2.pdf}
\caption{Example maps from two classes. Top: model 1 ($\Omega_m=0.31, \sigma_8=0.82$). Bottom: model 2 ($\Omega_m=0.26, \sigma_8=0.91$).
The initial conditions for these simulations were the same, so differences arise only due to different cosmological parameters. \nati{Tomek: Is this correct?}}
\label{fig:map_sample}
\end{figure*}

@@

We compare the performance of the spherical CNN to two simple benchmark
algorithms. The two baselines are based on building features with a) the Power
Spectral Densities or b) the histogram of pixels in the maps. CITE~\cite{...}
After a normalization, those features are then classified use a Support Vector
Machine (SVM) with a linear kernel. Other kernels did provide better result
while having worse scaling properties. For fairness, we augmented the dataset
of the baseline algorithm in a similar way as for the network, i.e, with
created an "unlimited" number of samples by adding new random realizations of
the noise. We stop adding new samples in the classifier once the validation
error converge to a values. This process is depicted in
Figure~\ref{fig:hist_error_evolution}.

The maps are separated into a training and a testing set. The performance is
assessed by measuring the classification error on a test set. Once the
training maps are sliced into samples, we split them into a training samples
and validation samples. The validation samples is used to tune the
regularization parameter of the SVM using a standard cross-validation
technique\footnote{We also give a small advantage to the SVM classifiers by
using cross-validation on the testing set.}.

@@

We present in Figure~\ref{fig:results} the results for 5 different noise levels and $3$ different data sizes. The spherical neural network is always significantly better than the hand-crafted classifier.

The histogram features contains information about the distribution of pixels, but does not include any spatial information.
In contrast, both the PSD features and the CNN can exploit the spatial information.

@@

\begin{itemize}
	\item Spherical CNN has the advantage to have some geometry information => one of the reason better than histogram
	\item PSD...
\end{itemize}

\citep{patton2017cosmologicalconstraints}

@@

\label{sec:spherical_harmonics}
\begin{figure*}[!ht]
\centering
\includegraphics[width=0.95\textwidth]{figures/eigenvectors.pdf}
\caption{12 first graph Fourier eigenvectors.}
\label{fig:graph_harmonics}
\end{figure*}

@@

>>>>

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.48\textwidth]{figures/smooth_map_class_1.pdf}
\includegraphics[width=0.48\textwidth]{figures/smooth_map_class_2.pdf}
\caption{Example maps from two classes. Top: model 1 ($\Omega_m=0.31, \sigma_8=0.82$). Bottom: model 2 ($\Omega_m=0.26, \sigma_8=0.91$).
The initial conditions for these simulations were the same, so differences arise only due to different cosmological parameters. \nati{Tomek: Is this correct?}}
\label{fig:map_sample}
\end{figure*}

@@

We compare the performance of the spherical CNN to two simple benchmark
algorithms. The two baselines are based on building features with a) the Power
Spectral Densities or b) the histogram of pixels in the maps~\cite{patton2017cosmologicalconstraints}.
After a normalization, those features are then classified use a Support Vector
Machine (SVM) with a linear kernel. Other kernels did not provide better result
while having worse scaling properties. For fairness, we augmented the dataset
of the baseline algorithm in a similar way as for the network, i.e, with
created an "unlimited" number of samples by adding new random realizations of
the noise. We stop adding new samples in the classifier once the validation
error converge to a value. This process is detailled in Appendix~\ref{sec:dataset_augmentation}.

The maps are separated into a training and a testing set. The performance is
assessed by measuring the classification error on a test set. Once the
training maps are sliced into samples, we split them into a training samples
and validation samples. The validation samples is used to tune the
regularization parameter of the SVM using a standard cross-validation
technique\footnote{We also give a small edge to the SVM classifiers by
using cross-validation on the testing set.}.

@@

We present in Figure~\ref{fig:results} the results for 5 different noise
levels and $3$ different data sizes.  We observe that the PSD features, i.e.
one of the most used technique for model discrimination in cosmology is unable
to perform the desired task. The main reason is that the two different model
have very similar PSDs (see Figure~\ref{fig:psd_sigma3}).   As a result, one
need to use other statistics to solve the problem. In this case, we use
histogram features and observe a significant improvement over the PSD.  The
SCNN present the advantage to learn automatically specific features for the
task and also the usual Deep Learning performance bump.

We also note that the histogram features contains information about the
distribution of pixels, but does not include any spatial information. In
contrast, both the PSD features and the SCNN can exploit the spatial
information. Without this scructure, a non scallable fully connected neural network approach would have to be used.

@@

@@

\label{sec:spherical_harmonics}
\begin{figure*}[!htb]
\centering
\includegraphics[width=0.95\textwidth]{figures/eigenvectors.pdf}
\caption{12 first graph Fourier eigenvectors.}
\label{fig:graph_harmonics}
\end{figure*}

@@

\section{Augementation of the dataset for the SVM classifier}
\label{sec:dataset_augmentation} The original number of samples $480$, $1920$,
$7680$ respectively for order $o=1,2,4$ is considered rather low for usual
deep learning technique, especially when one desire to build a noise robust
classifier. As a result, we use a classical dataset augmentation technique.
Before entering the network for training each sample is added a random
realization of Gaussian noise. The effect is twofold. First, the network never
ends up seing the same sample and is less likely to overfit the training set.
Second the trained network will be robust to Gaussian noise.

When comparing the network, we want to ensure that the SVM classifier has
access to the same amount of data that the network, i.e, potentially an
infinite number of sample. Hence, we fit different linear SVM classifier using
various amount of training set size until we oberve experimentally that
increasing the amount of data does not improve the classifier. The validation
set remain the same and is used for cross-validation to tune the L2
penalization parameter of SVM. Eventually, we evaluate the final classifier on
the testing set. An example of the evolution of the training and valdication
error is show in Figure~\ref{fig:hist_error_evolution}.
